MG7: Configurable and scalable 16S data analysis

### 2016-08-18 docs/mg7-paper/out/paper.tex

The starting point for our 16S reference database is RNA Central
\citep{rnacentral2014rnacentral},
\href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
5}. RNACentral was chosen being the most up to date, comprehensive RNA
sequence repository, including among others all RNA data from Silva,
GreenGenes, RDP, ENA, and RefSeq. First we take those sequences which

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  are annotated as being of \texttt{rRNA} type\footnote{We are aware of
    the existence of a gene annotation corresponding to 16S in
    RNACentra, that we are \textbf{not using} due to a significant
    amount of 16S sequences lacking it}
\item
  their length is at least \texttt{1300}\footnote{Note that 16S
    sequences are sometimes part of an entry corresponding to whole
    small subunits; that's why we do not set a maximum length threshold}
\item
  have at least one taxonomic assignment to a descendant of
  \emph{Bacteria} or \emph{Archaea}
\item
  their lineage does not contain a set of taxa deemed
  uninformative\footnote{for example: ``unclassified Bacteria
    (miscellaneous)'' or ``unclassified''. Assigning a read to an
    ``unclassified'' taxon defeats the first and foremost goal of
    taxonomic profiling: \emph{classification}.}
\end{enumerate}

@@

All these steps are automated, and integrated with every release of our
reference database. The whole process can be repeated and adapted to
other subsets of RNACentral; we do this for an 18S reference database.

>>>>

The starting point for our 16S reference database is RNA Central
\citep{rnacentral2014rnacentral},
\href{http://blog.rnacentral.org/2016/03/rnacentral-release-5.html}{version
5}. RNAcentral was chosen being the most up to date, comprehensive RNA
sequence repository, including among others all RNA data from Silva,
GreenGenes, RDP, ENA, and RefSeq. First we take those sequences which

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  are annotated as being of \texttt{rRNA} type\footnote{We are aware of
    the existence of a gene annotation corresponding to 16S in
    RNAcentral, that we are \textbf{not using} due to a significant
    amount of 16S sequences lacking it}
\item
  their length is at least \texttt{1300}\footnote{Note that 16S
    sequences are sometimes part of an entry corresponding to whole
    small subunits; that's why we do not set a maximum length threshold}
\item
  have at least one taxonomic assignment to a descendant of
  \emph{Bacteria} or \emph{Archaea}
\item
  their lineage does not contain a set of taxa deemed
  uninformative\footnote{for example: ``unclassified Bacteria
    (miscellaneous)'' or ``unclassified''. Assigning a read to an
    ``unclassified'' taxon defeats the first and foremost goal of
    taxonomic profiling: \emph{classification}.}
\end{enumerate}

@@

All these steps are automated, and integrated with every release of our
reference database. The whole process can be repeated and adapted to
other subsets of RNAcentral; we do this for an 18S reference database.
